\vspace{1.0cm} \LARGE \bf fluB

### 2013-07-24 fluB.tex
Minor changes to text, new figure.
Adding a figure showing the total time spent in "pure" vs "mixed"
PB1-PB2-HA.

\subsection*{Selection maintains the diversity in PB1-PB2-HA complexes of Victoria and Yamagata lineages}
Rather surprisingly, 3 segments have maintained a ratio of Vic and Yam lineages very close to 0.5 (PB1, PB2 and HA) (see Figure \ref{lineageRatios}) and have preserved the diversity that has accumulated since the split of the two lineages (see Figure \ref{tmrcaOT}).
Our results suggest that this is due to the 3 segments co-reassorting (comment - refer to posterior distributions of reassortant dates plot).
We find evidence that this co-reassortment pattern between PB1, PB2 and HA is not due to skewed reassortment (Marhsall \textit{et al.} \cite{marshall2013} have shown that reassortment efficiency is high in influenza A), but rather due to viruses bearing mixed-lineage PB1-PB2-HA complex having lower fitness than viruses with PB1-PB2-HA complexes derived entirely from Yam or Vic lineages (comment - refer to stratified REPU plot).
Not only are events that split up pure-lineage PB1-PB2-HA complexes rare, but they also have poor persistence and trees have spent less than 10\% of the time in the mixed-lineage PB1-PB2-HA complex state, suggesting that they are selected against (comment - make plot of that).

@@

\subsection*{Segments vary in their abilities to reassort and success as part of reassortant viruses}
A rather striking pattern in influenza B evolution is the fact that segments reassorted at the same time can have dramatically different fates.
For example, PA, NP, MP and NS segments of Yamagata lineage were reassorted into Victoria lineage PB1-PB2-HA background in 1994, but of those 4 segments only the NP lineage persisted for a considerable amount of time, being replaced by another NP lineage in 2006, whereas PA was replaced in 2000 and MP together with NS were replaced by different lineages in 2004.
This suggests that most non-PB1, PB2 or HA segments have the capacity to evolve somewhat independently of the rest of the segments present in the same virion.

@@

\subsection*{Implications for influenza B surveillance and prevention}
The findings presented here have several implications.
On one hand, if the co-dependence between PB1, PB2 and HA segments is strong enough, influenza B viruses containing mixed-lineage PB1-PB2-HA complex detected by surveillance are unlikely to contribute to long-term evolution of influenza B viruses.
On the other hand, the dependence of the HA segment on PB1 and PB2 segments (or \textit{vice versa}) implies that there is a possibility that more subtle disease control measures could be developed and used against influenza B viruses, or, if similar co-adapted segment complexes were discovered in influenza A, seasonal influenza in general.
In addition, the extent of inter-lineage reassortment argues for more intensive surveillance of influenza B viruses, as sequences of a handful of segments will not necessarily tell the whole story, especially since influenza B viruses have been known to evolve counter-intuitively (\textit{e.g.} losing Victoria lineages of PA and NA segments despite their associations with the rest of the polymerase complex or role in antagonising HA activity, respectively).
Finally, this study also highlights the gaps in our knowledge about the molecular biology of influenza B viruses.

>>>>

\subsection*{Selection maintains the diversity in PB1-PB2-HA complexes of Victoria and Yamagata lineages}
Rather surprisingly, 3 segments have maintained a ratio of Vic and Yam lineages very close to 0.5 (PB1, PB2 and HA) (see Figure \ref{lineageRatios}) and have preserved the diversity that has accumulated since the split of the two lineages (see Figure \ref{tmrcaOT}).
Our results suggest that this is due to the 3 segments co-reassorting (comment - refer to posterior distributions of reassortant dates plot).
We find evidence that this co-reassortment pattern between PB1, PB2 and HA is not due to skewed reassortment (Marhsall \textit{et al.} \cite{marshall2013} have shown that reassortment efficiency is high in influenza A), but rather due to viruses bearing mixed-lineage PB1-PB2-HA complex having lower fitness than viruses with PB1-PB2-HA complexes derived entirely from Yam or Vic lineages (comment - refer to stratified REPU plot).
Not only are events that split up pure-lineage PB1-PB2-HA complexes rare, but they also have poor persistence and trees have spent less than 10\% of the time in the mixed-lineage PB1-PB2-HA complex state (Figure \ref{stateTime}), suggesting that they are selected against (comment - make plot of that).

(comment - PB1, PB2 and HA are missing extra years in the mixed-lineage bar due to invisible reassortments e.g. VVY is invisible in HA tree, YVV in PB1 and YVY in PB2; solution: remove PB1, PB2 and HA bars or make it clear in the text)
\begin{figure}[h]
	\centering		
	\includegraphics[width=0.95\textwidth]{figures/InfB_stateTime.png}
	\caption{\textbf{Total time spent in "pure" vs "mixed" PB1-PB2-HA}}
	\label{stateTime}
\end{figure}

@@

\subsection*{Segments vary in their abilities to reassort and success as part of reassortant viruses}
A rather striking pattern in influenza B evolution is the fact that segments reassorted at the same time can have dramatically different fates.
For example, PA, NP, MP and NS segments of Yamagata lineage were reassorted into Victoria lineage PB1-PB2-HA background in 1994, but of those 4 segments only the NP lineage persisted for a considerable amount of time, being replaced by another NP lineage in 2006, whereas PA was replaced in 2000 and MP together with NS were replaced by different lineages in 2004.
This suggests that most non-PB1, PB2 or HA segments have the capacity to evolve somewhat independently of the rest of the genome.

@@

\subsection*{Implications for influenza B surveillance and prevention}
The findings presented here have several implications.
On one hand, if the co-dependence between PB1, PB2 and HA segments is strong enough, influenza B viruses containing mixed-lineage PB1-PB2-HA complex detected by surveillance are unlikely to contribute to long-term evolution of influenza B viruses.
On the other hand, the dependence of the HA segment on PB1 and PB2 segments (or \textit{vice versa}) implies that there is a possibility that more subtle disease control measures could be developed and used against influenza B viruses, or, if similar co-adapted segment complexes were discovered in influenza A, seasonal influenza in general.
In addition, the extent of inter-lineage reassortment argues for more intensive surveillance of influenza B viruses, as sequences of a handful of segments will not necessarily tell the whole story, especially since influenza B viruses have been known to evolve counter-intuitively (\textit{e.g.} losing Victoria lineages of PA and NA segments despite associations with the rest of the polymerase complex or antagonising HA activity, respectively).
Finally, this study also highlights the gaps in our knowledge about the molecular biology of influenza B viruses.
